---
title: "zchu_Final"
author: "Dolores Chu"
date: "12/4/2019"
output: html_document
---

##### Load and examine data.
```{r}
data <- read.csv("BathSoap.csv")
str(data)
head(data)
colnames(data)
```

##### Tranform some variables from factors to numeric.
```{r}
data[,20] <- (as.numeric(gsub("%", "", (as.character(data[, 20])))))/100
data[,21] <- (as.numeric(gsub("%", "", (as.character(data[, 21])))))/100
data[,22] <- (as.numeric(gsub("%", "", (as.character(data[, 22])))))/100
data[,23] <- (as.numeric(gsub("%", "", (as.character(data[, 23])))))/100
data[,24] <- (as.numeric(gsub("%", "", (as.character(data[, 24])))))/100
data[,25] <- (as.numeric(gsub("%", "", (as.character(data[, 25])))))/100
data[,26] <- (as.numeric(gsub("%", "", (as.character(data[, 26])))))/100
data[,27] <- (as.numeric(gsub("%", "", (as.character(data[, 27])))))/100
data[,28] <- (as.numeric(gsub("%", "", (as.character(data[, 28])))))/100
data[,29] <- (as.numeric(gsub("%", "", (as.character(data[, 29])))))/100
data[,30] <- (as.numeric(gsub("%", "", (as.character(data[, 30])))))/100
data[,31] <- (as.numeric(gsub("%", "", (as.character(data[, 31])))))/100
data[,32] <- (as.numeric(gsub("%", "", (as.character(data[, 32])))))/100
data[,33] <- (as.numeric(gsub("%", "", (as.character(data[, 33])))))/100
data[,34] <- (as.numeric(gsub("%", "", (as.character(data[, 34])))))/100
data[,35] <- (as.numeric(gsub("%", "", (as.character(data[, 35])))))/100
data[,36] <- (as.numeric(gsub("%", "", (as.character(data[, 36])))))/100
data[,37] <- (as.numeric(gsub("%", "", (as.character(data[, 37])))))/100
data[,38] <- (as.numeric(gsub("%", "", (as.character(data[, 38])))))/100
data[,39] <- (as.numeric(gsub("%", "", (as.character(data[, 39])))))/100
data[,40] <- (as.numeric(gsub("%", "", (as.character(data[, 40])))))/100
data[,41] <- (as.numeric(gsub("%", "", (as.character(data[, 41])))))/100
data[,42] <- (as.numeric(gsub("%", "", (as.character(data[, 42])))))/100
data[,43] <- (as.numeric(gsub("%", "", (as.character(data[, 43])))))/100
data[,44] <- (as.numeric(gsub("%", "", (as.character(data[, 44])))))/100
data[,45] <- (as.numeric(gsub("%", "", (as.character(data[, 45])))))/100
data[,46] <- (as.numeric(gsub("%", "", (as.character(data[, 46])))))/100
```

##### For the brand purchase percentage variables, pick the highest one for each observation and assign it to a new variabel called "Max.Br.Cd". Then remove the individual brand purchase percentage variables. Also remove the demographic variables. 
```{r}
data$Max.Br.Cd <- apply(data[,23:31], 1, max)
data <- data[,-c(1:11,23:31)]
```

##### Normalize continuous variables.
```{r}
library(caret)
norm_model<-preProcess(data,method = "range")
data_norm <- predict(norm_model,data)
```


##### Selecting optimal k. It seems like k = 2 is the best choice.
```{r}
library(factoextra)
distance <- get_dist(data_norm)
set.seed(123)
fviz_nbclust(data_norm, kmeans, method = "wss")
fviz_nbclust(data_norm, kmeans, method = "silhouette")
```

##### Apply k-means algorithm with k = 2.
```{r}
k2 <- kmeans(data_norm,centers = 2, nstart = 25)
k2$centers
k2$size
fviz_cluster(k2, data = data_norm)
```

##### Analyze the clusters
```{r}
data$cluster <- k2$cluster
Cluster_1 <- data[data$cluster == 1, ]
Cluster_2 <- data[data$cluster == 2, ]
summary(Cluster_1)
summary(Cluster_2)
```

##### To develop a model to classify future data, I define my target segment as customers who have lower brand loyalty and higher purchasing power (Cluster 1). I assume the model will be used by businesses that are trying to gain new customers.
##### I use the k-nn algorithm to buil the model, k=7.
```{r}
set.seed(123)
library(e1071)
data_norm$cluster <- k2$cluster
data_norm$cluster <- as.factor(data_norm$cluster)
search_grid <- expand.grid(k=c(1:10))
model <- train(cluster~., data = data_norm, method = "knn", tuneGrid=search_grid)
model
```
