---
title: "Zchu_HW2_FlightDelays"
author: "Z Chu"
date: "10/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Divide the data into 60% training and 40% validation
```{r}
#Read in data. Load library.
data<-read.csv("FlightDelays.csv")
library("caret")

#Transform int variables into factors. I transformed day of the week, date of the month, time and also weather variables. 
data$CRS_DEP_TIME <- as.factor(data$CRS_DEP_TIME)
data$DEP_TIME <- as.factor(data$DEP_TIME)
data$Weather <- as.factor(data$Weather)
data$DAY_WEEK <- as.factor(data$DAY_WEEK)
data$DAY_OF_MONTH <- as.factor(data$DAY_OF_MONTH)

#Partition the data into 60% training and 40% validation.
set.seed(123)
Train_Data_Index <- createDataPartition(data$Flight.Status, p = 0.6, list = FALSE)
Train_Data <- data[Train_Data_Index, ]
Validation_Data <- data[-Train_Data_Index, ]
```

2. Run the Naive Bayes model to predict whether the flight is delayed or not. Use only categorical variables for the predictor variables. Note that Week and Time variables need to be recoded as factors
```{r}
#Load the e1071 library and the ISLR library.
library(e1071)
library(ISLR)

#Remove the integer variables
data<-data[, -c(5, 7)]

#Build the naive Bayes classifer using training data
nb_model <- naiveBayes(Flight.Status~CRS_DEP_TIME+CARRIER+DEP_TIME+DEST+FL_DATE+ORIGIN+Weather+DAY_WEEK+DAY_OF_MONTH+TAIL_NUM, data = data)

#Use the model built above to predict the validation set
Predicted_Validation_Labels <- predict(nb_model, Validation_Data)
head(Predicted_Validation_Labels)
```

3. Output both a counts table and a proportion table outlining how many and what proportion of flights were delayed and on-time at each of the three airports.
```{r}
library(dplyr)
Validation_Data <- cbind(Validation_Data, Predicted_Validation_Labels)
#Creating a count table for number of delayed or on time flights at each of the three airports. 
Validation_Data %>% group_by(ORIGIN) %>% summarize(On_Time = sum(Predicted_Validation_Labels == "ontime"), Delayed = sum(Predicted_Validation_Labels == "delayed"))

#Creating a proportion table for delayed or on time flights at each of the three airports. 
Validation_Data %>% group_by(ORIGIN) %>% summarize(On_Time_Percentage = sum(Predicted_Validation_Labels == "ontime")/n(), Delayed_Percentage = sum(Predicted_Validation_Labels == "delayed")/n())
```

4. Output the confusion matrix and ROC for the validation data
```{r}
#Load needd libraries.
library(gmodels)
library(pROC)

#Create the confusion matrix.
CrossTable(x=Validation_Data$Flight.Status, y=Predicted_Validation_Labels, prop.chisq = FALSE)

#Create the ROC curve.
#First, change the model slightly to return the probablity of the prediction.
Predicted_Validation_Labels <- predict(nb_model, Validation_Data, type = "raw")
head(Predicted_Validation_Labels)
roc(Validation_Data$Flight.Status, Predicted_Validation_Labels[, 1])
plot.roc(Validation_Data$Flight.Status, Predicted_Validation_Labels[, 1])
```