---
title: "Machine Learining_Project 1"
output: html_notebook
---

Read in file.Intall caret.
```{r}
UBdata<-read.csv("UniversalBank.csv")
head(UBdata)
summary(UBdata)

install.packages("caret")
library(caret)
```

Transform categorical predictors with more than two categories into dummy variables. In this case, Education.
```{r}
UBdata$Education<-as.factor(UBdata$Education)
dummy_model<-dummyVars(~Education,data=UBdata)
head(predict(dummy_model, UBdata))
Education_dummy<-predict(dummy_model, UBdata)
Education_dummy
UBdata<-cbind(UBdata, predict(dummy_model, UBdata))
UBdata$Education<- NULL
UBdata
```

Nomralize Data
```{r}
Norm_UBdata<-preProcess(UBdata[, c(2, 3, 4, 6, 7, 8)], method = c("center", "scale"))
UBdata_normalized<- predict(Norm_UBdata,UBdata)
summary(UBdata_normalized)
```

Partition the data into training (60%) and validation (40%) sets.
```{r}
Train_Index<-createDataPartition(UBdata$Personal.Loan, p = 0.6, list=FALSE)
Train_Data<-UBdata[Train_Index,]
Test_Data<-UBdata[-Train_Index,]
```

Examine that that 2 sets have similar distribution in personal loan acceptance rate
```{r}
summary(Train_Data$Personal.Loan)
summary(Test_Data$Personal.Loan)
```

Train the model using the training set.
```{r}
set.seed(123)
UBdata_normalized$Personal.Loan<- as.factor(UBdata_normalized$Personal.Loan)
model<-train(Personal.Loan~Age+Experience+Income+Family+CCAvg+Mortgage+Securities.Account+CD.Account+Online+CreditCard+Education.1+Education.2+Education.3, data = UBdata_normalized, method = "knn")
model
```
Use the model with k = 7 to predict the given case
```{r}
install.packages("class")
library(class)
library(ISLR)
Test_Customer<- data.frame("Age" = 40, "Experience" = 10, "Income" = 84, "Family" = 2, "CCAvg" = 2, "Education_1" = 0, "Education_2" = 1, "Education_3" = 0, "Mortgage" = 0, "Securities Account" = 0, "CD Account" = 0, "Online" = 1, "Credit Card" = 1)

Train_Predictors <- Train_Data[, c(2:4, 6:8, 10:16)]
Customer_Predictors <- Test_Customer
Train_Labels <- Train_Data[, 9]
Predicted_customer_labels <- knn(Train_Predictors, Customer_Predictors, cl = Train_Labels, k = 7)
Predicted_customer_labels
#According to the model, this customer would be classified as unsuccessful (loan not accepted).
```

2. What is a choice of k that balances between overfitting and ignoring the predictor information?
k = 7.

3. Show the confusion matrix for the validation data that results from using the best k.
```{r}
Test_Predictors <- Test_Data[, c(2:4, 6:8, 10:16)]
Test_Labels <- Test_Data[, 9]
Predicted_test_labels <- knn(Train_Predictors, Test_Predictors, cl = Train_Labels, k = 7)

install.packages("gmodels")
library(gmodels)

CrossTable(x=Test_Labels, y=Predicted_test_labels, prop.chisq = FALSE)
```


4. Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.
```{r}
Customer_2_Predictors <- data.frame("Age" = 40, "Experience" = 10, "Income" = 84,
"Family" = 2, "CCAvg" = 2, "Education_1" = 0, "Education_2" = 1, "Education_3" = 0,
"Mortgage" = 0, "Securities Account" = 0, "CD Account" = 0, "Online" = 1., "Credit
Card" = 1)
Predicted_customer_2_label <- knn(Train_Predictors, Customer_2_Predictors, cl = Train_Labels, k = 7)
Predicted_customer_2_label
```



5. Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason.
```{r}
Train_Index = createDataPartition(UBdata$Personal.Loan, p = 0.5, list = FALSE)
Train_Data = UBdata[Train_Index,]
Non_Train_Data = UBdata[-Train_Index,]
Validation_Index = createDataPartition(Non_Train_Data$Personal.Loan, p = 0.6, list = FALSE)
Validation_Data = Non_Train_Data[Validation_Index, ]
Test_Data = Non_Train_Data[-Validation_Index,]
summary(Train_Data$Personal.Loan)
summary(Validation_Data$Personal.Loan)
summary(Test_Data$Personal.Loan)
set.seed(123)
Train_Predictors <- Train_Data[, c(2:4, 6:8, 10:16)]
Validation_Predictors <- Validation_Data[, c(2:4, 6:8, 10:16)]
Test_Predictors <- Test_Data[, c(2:4, 6:8, 10:16)]
Train_Labels <- Train_Data[, 9]
Validation_Labels <- Validation_Data[, 9]
Test_Labels <- Test_Data[, 9]
Predicted_Validation_Labels <- knn(Train_Predictors, Validation_Predictors, cl = Train_Labels, k = 7)
Predicted_Test_Labels <- knn(Train_Predictors, Test_Predictors, cl = Train_Labels, k = 7)
head(Predicted_Test_Labels)
CrossTable(x = Validation_Labels, y = Predicted_Validation_Labels)
CrossTable(x = Test_Labels, y = Predicted_Test_Labels)
```


